---
title: "HW04 Part 1: Complete Chapter 3"
author: "Taylor Simmonds"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

Loaded mpg data frame and found its dimensions. Created scatterplot comparing hwy/cyl and class/drv.
```{r}
ggplot2::mpg
dim(mpg)
?mpg
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y =cyl))
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y=drv))
```


#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
A data frame containing variables in the mpg dataset.

**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.
243 rows and 11 colums.

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.
drv describes the style of driving for f = front, r=rear wheel drive and 4=4wd 

**4:** Make a scatterplot of `hwy` vs `cyl`.


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?
The driving style of the car and the type of car show no significant data trends when graphed together, these factors alone do not help in comparing fuel economy across various cars.


### Section 3.3 
Shows a scatterplot of displ and hwy
```{r}
?points
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=hwy, color = "blue"))
 
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=hwy), color="blue", pch=1, cex=2)
```
Display a scatterplot for drv/class
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x=drv, y=class), color="blue", pch=1, cex=2)
```
Applying two color aesthetics to the same variable
```{r}
ggplot2::data = mpg +
  geom_point(mapping = aes(x=displ, y=hwy), color="blue", color="red", pch=1, cex=2)
```
Applying an aesthetic to something other than a variable name.
```{r}
ggplot2::data = mpg +
  geom_point(mapping =aes(colour = displ < 5))
```

### Section 3.3: Aesthetic mappings

#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?
It needs to be written (mapping = aes(x=displ, y=hwy), color="blue"). 

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?
Model name, type of transmission, drv, fl, and class. When mpg? is ran a help window shows in the bottom right of the screen. 

**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.
The graphs with categorical variables do not show important correlations, while the continous variables show some sort of importance in the data.

**4:** What happens if you map the same variable to multiple aesthetics?

An error appears when multiple aesthetics are mapped to the same variable.

**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)
The stroke aesthetic allows for syombols to have a color and a fill, it works for shapes 21-24.


**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?
An error for incompatible methods appears, only specific colors can be assigned to "color", numerical values would not work.

### Section 3.5: Facets
Facet plot for mpg
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
Facet plot for drv and class.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv~ class)
```
Facet plotted with "."
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
Plot for displ and hwy with a facet for class 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
?facet_grid
```

#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.

This should split the plot into various subplots that each display a specific portion of the dataset.

**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
The empty cells show that those car styles do not come in rear wheel drive and that the 2 seater car does not come in 4 wheel or front wheel drive. This shows what style of driving is available for each car type. 


**3:** What plots does the following code make? What does `.` do?
"drv~." creates a subplot that does not properly split the displ variable across the graph and ".~cyl" creates a subplot that shows the displ and cyl variable. The "." tells the subplot that there should be no faceting on that specific variable. 


**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Facet would be a quick way to sort small amounts of data, but color would be easier and quicker to digest and understand larger datasets. 


**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?
nrow and ncol set the number of rows and columns respectively. Labeller and switch control the layout of the individual panels. "facet_grid" has rows/cols because instead of setting the number of rows/cols it sets the name of them through labeller.  

**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?
A facet forms a matrix of panels usually with two discrete variables and all combinations pf the variable that exists in the data, placing the variable with the more uniqure levels in the columns would make the graph easier to read.
### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.



**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.


**4:** What does the `se` argument to `geom_smooth()` do?


**5:** Will these two graphs look different? Why/why not?


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.


2: upper right.


3: middle left.


4: middle right.


5: lower left.


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 


**2:** What parameters to `geom_jitter()` control the amount of jittering?


**3:** Compare and contracts `geom_jitter()` with `geom_count()`.


**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


